From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-05-30 paper.tex
Review Livre section

Real-time rendering of volumetric data suffers from the curse of
dimensionality. The data size of regular grids increases cubically as the number
of dimensions increase. To render such data interactively, out-of-core (OOC)
storage using multi-resolution algorithms are commonly employed.

Livre is an interactive volume rendering engine available under a permissive
open source license. Our main contributions are: an state-of-the-art
implementation of an octree-based level-of-detail selection, a task parallel
rendering pipeline, a multi-GPU based parallel rendering engine, and an easily
extensible renderer through the use of plugin data sources. Furthermore we
propose to create volume representations of different input data using
on-the-fly, interactive transformations of the data, skipping costly
preprocessing steps.

Our system brings together state-of-the art algorithms to create a volume
rendering engine capable of handling extremely high-resolution volumes using a
high degree of parallelism, both on a single system as well as in a distributed
cluster. We employ a GPU-based ray casting algorithm to compute the radiance
absorption of the given volumetric data. The computation is executed per pixel
on the pixel shader hardware of the GPU.

In our out-of-core data access layer, multi-resolution data is represented as an
octree data structure. This representation accelerates the selection of the
proper level-of-detail and to track the status of the LODs (in CPU memory, in
GPU memory, not loaded). While rendering, view-based LOD selection is performed
using the on the screen-space-error (SSE)~\cite{guthe2004} technique. According
to the requested maximum SSE, the selected LODs represent the given volume the
best on the screen.

The creation of volume bricks, their upload to the GPU and the rendering are
executed in separate tasks. These tasks run asynchronously, that is, they do not
block each other. This is particularly important for the render thread, which
has to react to user input as quickly as possible to reduce input latency, and
renders at interactive frame rates. The data loader thread can take seconds to
generate the brick, for example in the field voxelization data source when
sampling millions of events.

Livre uses a plugin mechanism to access the volume data. In this mechanism, data
sources are implemented as shared libraries and are loaded on application
startup based on the URI of the input data. Loading the shared libraries on
runtime allows to add third-party data loaders to an existing Livre application
by simply placing them into the shared library directory. Data sources have to
provide the requested volume bricks, that is, there is no defined file format or
even requirement to read the input data from a file system. This flexibility of
the plugin approach lead to novel volume rendering use cases, where volume
representations are created on the fly from different input data sets.

@@

We have presented a modular visualization architecture for large data
visualization over a wide range of use cases, glued together by a modern and
easy to use messaging infrastructure. This ecosystem allows us to flexibly
support novel use cases, while pushing novel visualization capabilities,
providing Blue Brain with a competitive visualization infrastructure. Messaging
and remote APIs not only surprised us in their versatility and ease of
integration with other tools, but have still a significant potential for future
exploration of classical visualization software. Interactive raytracing is the
rendering algorithm for us, and Brayns is becoming the integration point for our
domain-specific visual applications and algorithms. Tiled display walls are
affordable for a large set of institutions, and coupled with our open source
TIDE software create new ways of truely collaborative work. TIDE, together with
cheaper visualization hardware, will evolve in the future towards a seamless
integration of immersive visualization.

>>>>

Livre is an interactive volume rendering engine available under a permissive
open source license. Our main contributions are: a state-of-the-art
implementation of an octree-based level-of-detail (LOD) selection, a task
parallel rendering pipeline, a multi-GPU parallel rendering engine, and an
easily extensible renderer through the use of plugin data sources. Our system
brings together state-of-the art algorithms to create a volume rendering engine
capable of handling extremely high-resolution volumes using a high degree of
parallelism, both on a single system and in a distributed cluster. We employ a
GPU-based ray casting algorithm to compute the radiance absorption of the given
volumetric data. The computation is executed per pixel on the pixel shader
hardware of the GPU.

In our out-of-core data access layer, multi-resolution data is represented as an
octree data structure. This representation accelerates the selection of the
proper level-of-detail and to track the status of the LODs (in CPU memory, in
GPU memory, not loaded). While rendering, view-based LOD selection is performed
using the screen-space-error (SSE)~\cite{guthe2004} technique. \fig{fLivre}
shows the rendering of 300 GB volume with an illustration of the selected LOD
levels.

\begin{figure}[h!t]
    \includegraphics[width=.45\columnwidth]{images/livre}\hfill
    \includegraphics[width=.45\columnwidth]{images/livreLOD}
  \caption{\label{fLivre}Livre rendering a 300GB vasculature dataset (left) and
    the selected LODs (right)}
\end{figure}

The creation of volume bricks, their upload to the GPU and the rendering are
executed in separate tasks. These tasks run asynchronously, that is, rendering
is decoupled from data loading. Livre uses a plugin mechanism to access the
volume data, where data sources are implemented as shared libraries and are
loaded on application startup based on the URI of the input data. Data sources
only have to provide the requested volume bricks, that is, there is no defined
file format or even requirement to read the input data from a file system. This
flexibility of the plugin approach lead to novel volume rendering use cases,
where volume representations are created on the fly from different input data
sets, for example from simulation data.

@@

We presented a modular visualization architecture for large data visualization
over a wide range of use cases, glued together by a modern and easy to use
messaging infrastructure. This ecosystem allows us to flexibly support novel use
cases, while pushing novel visualization capabilities, providing Blue Brain with
a competitive visualization infrastructure. Messaging and remote APIs not only
surprised us in their versatility and ease of integration with other ecosystems,
but also have a significant potential for future exploration in classical
visualization software. Interactive raytracing is the future rendering algorithm
for us, and Brayns is becoming the integration point for our domain-specific
visual applications and algorithms. Tiled display walls are affordable for a
large set of institutions, and coupled with our open source TIDE software create
new ways of truely collaborative work. TIDE, together with cheaper visualization
hardware, will evolve in the future towards a seamless integration of immersive
visualization.
